---
title: "2d Bermudan Basket Put Demo"
output: html_document
---

This is a demo file to generate Figure 6 in paper "Adaptive Batching for Gaussian Process Surrogates with Application in Noisy Level Set Estimation". The plot shows the fitted exercise boundary with its 95\% credible interval (solid line and dashed line) obtained with Gaussian Process for two-dimensional basket put Bermudan option. Two batch heuristics, ABSUR and ADSA, are used to select the location of inputs and their replications, which are shown as the dots and their color/size.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ks)
library(fields) # for plotting purposes, use quilt.plot in 2D
library(mlOSP)
library(DiceKriging)
library(tgp)  # use lhs from there
library(randtoolbox)  # use sobol and halton QMC sequences
library(hetGP)
library(laGP)
library(ggplot2)
library(pander)
data("int_2d")
```

## Set up the model for Two-dim basket put with parameters in Table 4
```{r}
model2d <- list(x0 = rep(40,2),K=40,sigma=rep(0.2,2),r=0.06,div=0,T=1,dt=0.04,dim=2,sim.func=sim.gbm, payoff.func=put.payoff)
model2d$pilot.nsims <- 1000
model2d$look.ahead <- 1
model2d$cand.len <- 1000   # size of testing set m_0
model2d$max.lengthscale <- c(40,40)
model2d$min.lengthscale <- c(3,3)
model2d$tmse.eps <- 0

model2d$ucb.gamma <- 1.96

model2d$seq.design.size <- 100  # budget N = r_0 * k = 2000
model2d$batch.nrep <- 20  # initial replication r_0

model2d$init.size <- 20   # initial design size k_0
model2d$init.grid <- int_2d

model2d$tmse.eps <- 0
model2d$kernel.family <- "gauss"  # kernel function for Gaussian Process
model2d$ucb.gamma <- 1.96
model2d$update.freq <- 5   # number of sequential design steps to update the GP surrogate
nReps <- 2
model2d$r.cand <- c(20, 30,40,50,60, 80, 120, 160) # r_L
```

## Compare results for GP with ABSUR and ADSA
```{r testing-homtp2d, message=FALSE, warning=FALSE}
### GP + ADSA
set.seed(110)
model2d$batch.heuristic <- 'adsa'
model2d$ei.func <- "amcu"
oos.obj.adsa <- osp.seq.batch.design(model2d, method="trainkm")

### GP + ABSUR
set.seed(122)
model2d$batch.heuristic <- 'absur'
model2d$ei.func <- 'absur'
oos.obj.absur <- osp.seq.batch.design(model2d, method="trainkm")

### plot Figure 6
plt.2d.surf.batch(oos.obj.adsa$fit[[15]], oos.obj.absur$fit[[15]], oos.obj.adsa$batches[1:oos.obj.adsa$ndesigns[15] - 1, 15], oos.obj.absur$batches[1:oos.obj.absur$ndesigns[15] - 1, 15], "ADSA", "ABSUR", x=seq(25,50,len=201),y = seq(25, 50,len=201))
```









